---
title: "ESM244 Lab 1"
author: "Allison Hacker"
date: "1/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(here)
library(sf)
library(tmap)
```

### Read in SF trees data

```{r}

sf_trees <- read_csv(here("data","sf_trees","sf_trees.csv"))

```

### Basic wrangling reminders

Data wrangling and summary statistics using functions in the `dplyr` package

Find the top 5 highest observations of trees by legal status, wrangle, and make a graph.
```{r}

### Wrangling

top_5_status <- sf_trees %>%
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>%
  slice_max(tree_count, n=5)

### Make a graph with ggplot

ggplot(data = top_5_status, aes(x = fct_reorder(legal_status,tree_count), y = tree_count))+
  geom_col()+
  labs(x = "Legal Status", y = "Tree Count")+
  coord_flip()+
  theme_minimal()
```

### More data wrangling examples

Only want to keep observations for Blackwood Acacia trees, plot using lat and long

```{r}

blackwood_acacia <- sf_trees %>%
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, latitude, longitude)

ggplot(data= blackwood_acacia, aes(x = latitude, y = longitude))+
  geom_point()
```

### tidyr:: separate() and unite () functions

Useful for separating columns
```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")

sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_|_|_")
```

### Make some maps of trees in SF

Use `st_as_sf()` converts lat and long to spatial coordinates

Spatial data is sticky

```{r}
blackwood_acacia_sp <- blackwood_acacia %>% 
  drop_na(latitude, longitude) %>% 
  st_as_sf(coords = c("longitude", "latitude"))

# needs coordinate reference system

st_crs(blackwood_acacia_sp) = 4326

ggplot(data = blackwood_acacia_sp)+
  geom_sf(color = "darkgreen")

# read in SF roads shapefile

sf_map <- read_sf(here("data", "sf_map", "tl_2017_06075_roads.shp"))

st_transform(sf_map, 4326)

ggplot(data = sf_map)+
  geom_sf()

# combine blackwood acacia trees and sf roads map

ggplot()+
  geom_sf(data = sf_map, size = 0.1, color = "darkgray")+
  geom_sf(data = blackwood_acacia_sp, color = "red", size = 0.4)+
  theme_void()


```
 ### Make interactive map using `tmap`
 
```{r}
# set mode to interactive map
 tmap_mode("view")

tm_shape(blackwood_acacia_sp)+
  tm_dots()
```
 

